Global Optimization of an Analog Method by Means of Genetic Algorithms

### 2016-09-22 Genetic algorithms - MWR/template.tex
Elsevier english edition.

\abstract{Analog methods are based on a statistical relationship between synoptic meteorological variables (predictors) and local weather (predictand, which are to be predicted), which is defined by several parameters that are often calibrated by means of a semi-automatic sequential procedure. This calibration approach is fast, but has strong limitations: it is made of successive steps and thus cannot handle all parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows (hour(s) of the day) on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely Genetic Algorithms, which can optimize jointly all parameters of the method and can get closer to a global optimum, by taking into account the parameters dependencies. Moreover, it can objectively calibrate parameters that were previously manually assessed, and can take into account new degrees of freedom.
Genetic Algorithms need however to be tailored to the addressed problem. Multiple combinations of algorithms were assessed, and new ones were developed (e.g., the \textit{chromosome of adaptive search radius} which was found to be very robust), in order to make recommendations for the use of Genetic Algorithms to optimize several variants of analog methods. A global optimization approach opens new perspectives for the improvement of analog methods, and their application to new regions or to new predictands.
}

@@

Analog methods (AMs) rely on the hypothesis that similar situations in terms of the atmospheric circulation are likely to lead to similar local weather \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. They consist in sampling a certain number of past situations based on different synoptic-scale meteorological variables (predictors) in order to build a probabilistic prediction for a local weather variable of interest (predictand). A common usage of AMs is for operational precipitation forecasting \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015b, BenDaoud2016}, or more recently for precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. However, AMs or equivalent are also used to predict temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

Even though the method is rather simple, it contains several parameters to optimize, such as the choice of the predictor variables, the pressure levels and temporal windows (hour(s) of the day) on which the predictors are compared, the spatial domains used for the comparison, as well as the analogy criteria (distance measure), and finally, the number of analog situations to retain.

AMs need to be adapted to every new region considered, because the leading meteorological influences may be location specific. Even the selection of the pressure levels and the temporal windows should be reconsidered, if not the predictor variable itself. Thus, before being applied in a forecasting or downscaling context, AMs must be calibrated for the given region, which is done here in the Perfect Prognosis framework \citep{Klein1963}, like the majority of AMs calibration procedures for precipitation prediction cited previously \cite[with the exception of][]{Hamill2006,Hamill2015b}.

A common approach to optimize the method is by means of a semi-automatic sequential calibration procedure, developed by \citet{Bontron2004}, also described in \citet{BenDaoud2016}, and extended by \citet{Radanovics2013}. It determines some parameters of the method, sequentially for each consecutive analogy level (e.g., on the atmospheric circulation or on a moisture index). It starts by a manual selection of the meteorological variables (e.g., geopotential height, relative humidity, etc.), the pressure levels, the temporal windows, and the initial analog numbers. Then, the spatial window on which the predictors are compared is optimized by an iterative growth of the domain, and the number of analogs is finally reassessed. A successive level of analogy can then be added and its spatial window optimized. Parameters of the preceding levels of analogy are not reassessed, except for the number of analog situations to keep.

The sequential calibration procedure thus allows optimizing a limited number of parameters (spatial windows and analog numbers), but the selection of predictor variables, pressure levels and temporal windows have still to be made manually. Testing multiple combinations of these is very combinatorial and becomes quickly cumbersome, especially when considering multiple predictors within the same level of analogy. Thus, optimizing AMs by means of the sequential technique is laborious when little knowledge on the predictor-predictand relationship or the leading meteorological influences is available. Indeed, many combinations of parameters (predictor variables, pressure levels, temporal windows) have to be assessed. Moreover, proceeding to the optimization sequentially ignores potential dependencies between the parameters of the method, may they be within a single level of analogy or between them, which could lead to another configuration if the parameters were calibrated together. Thus, the risk of ending in a local optimum is high and can not be avoided.

When creating the sequential calibration procedure, \citet{Bontron2004} was aware of the problem of dependencies between parameters and wrote: ''\textit{We perceive here the combinatorial aspect of our problem: variables and spatial windows are not independent. We will present our results by first searching the best variable (e.g., selection of the pressure level and the temporal window for the geopotential height) on a chosen spatial window, and next, the best window for the chosen variable. However, even by repeating the process, are we sure to obtain the optimal combination?}''. And later in his work: ''\textit{Our approach, which is again to vary the parameters one by one -- the others being fixed in a more or less arbitrary manner -- may therefore not exactly lead us to the optimal solution}''. \citet{Bliefernicht2010} has also faced the combinatorial issue of AMs parameters and concluded that one needs to be an expert to have a sense of their respective influence, sensitivity and nonlinear interactions. \citet{BenDaoud2010}, when calibrating an AM, also stated that ''\textit{the combinatory aspect related to the calibration was found to be too high for all the parameters to be calibrated simultaneously}''. Simultaneous calibration of all parameters has never been undertaken so far.

Another optimization strategy is proposed by \citet{Junk2015}, and allows for an automatic calibration of weights applied to the different predictors when processing the analogy criteria (distance function). Their strategy consists in a brute-force assessment of all possible combinations. This approach is possible in their implementation, as predictors are considered at a unique point (interpolated to the location of interest), at fixed hours, and at preselected pressure levels, leaving only the weights to be optimized. In the presently used AM (described in section \ref{sec:am}), the number of parameters to optimize makes it impossible to proceed to a brute-force strategy.

In order to overcome these limitations, two optimization techniques were assessed. First, \citet{Horton2012a} assessed the ability of the \citet{Nelder1965a} method based on a simplex algorithm. This technique did not provide satisfying results and failed at converging toward a unique solution. The parameter space of AMs can be very complex and is inappropriate for a linear optimization technique. Thus, global optimization techniques are likely to be necessary in order to calibrate most AM variants, as they can optimize all parameters of all analogy levels simultaneously. In addition, they can overcome the systematic manual assessments of all pressure levels and temporal windows. Finally, they allow testing new degrees of freedom in AMs. The relevance of Genetic Algorithms (GAs) is presented here, which does not exclude that other global optimization techniques could also work. If using GAs to optimize AMs is computationally intensive, once an AM is calibrated, its use in real-time operations or climate downscaling is very fast and lightweight.

This paper specifically describes how GAs are to be used in order to successfully optimize several AMs. Indeed, GAs variants are numerous and always need to be tailored to the addressed problem. This requires intensive and systematic comparisons of operators and options in order to identify the key factors leading the optimization and the respective sensitivity of the options. Such analyzes are presented here and result in recommendations for the use of GAs to optimize several AM implementations. The in-depth analysis of the benefit brought by such an approach on a specific case study is the topic of another coming paper.

The document starts by presenting the case study area and the data (section \ref{sec:case_study_data}), the considered AM variants (section \ref{sec:am}), and the assessed GAs options (section \ref{sec:gas}). The comparative analyzes of these options are presented in section \ref{sec:assessment}, which results in recommendations formulated in section \ref{sec:use} and conclusions in section \ref{sec:conclusions}.

@@

The study area is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l.\ and the area is 5524~km$^{2}$.  Based on different climatological analyzes \cite[see][for the details]{Horton2012a}, the gauging stations in the catchment were clustered in 10~subregions (Fig.\ \ref{fig:map}).

This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project that aims at providing real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. AMs are used to provide real-time probabilistic precipitation forecasts, based on NWP outputs.

@@

AMs rely on two types of data: predictors, that are meteorological variables describing the state of the atmosphere at a synoptic scale, and the predictand, which is the local weather time series one wants to predict.

When working in the Perfect Prognosis framework \citep{Klein1963}, the meteorological archive from which the predictors are extracted is usually a reanalysis dataset. Oppositely, other applications of AMs for wind forecasting are based on a Model Output Statistics context \citep[MOS, see][]{Glahn1972}, and thus use forecast archives or reforecast products \citep[e.g.,][]{DelleMonache2013, DelleMonache2011, Alessandrini2015, Junk2015, Junk2015c}.

In the present study, the NCEP-NCAR reanalysis I dataset \citep[6-hourly, 17 pressure levels at a resolution of 2.5\degree, see][]{Kalnay1996} was used, but it could have been any other reanalysis dataset. This dataset is relatively old, even though still widely used, and better results may be expected with more recent datasets. However, we can safely make the assumption that if the optimization technique works for this reanalysis dataset, it will also work on an improved one.

The predictand (which is to be predicted) was here the daily precipitation (0540~UTC to 0540~UTC the following day) measured at the MeteoSwiss' stations network in the catchment of interest. The time series from every available gauging station were averaged over subregions of approximately 500~km$^{2}$ \citep[see][for the details]{Horton2012a} in order to smooth local effects \citep{Obled2002, Marty2012}. The time series need to be split into a calibration and an independent validation periods.

@@

AMs are based on the principle that two similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. Thus, they consist in searching a certain number of past situations in a meteorological archive that present similar properties for chosen predictors, according to an analogy criteria, in order to extract the observed values of the local weather variable of interest from another archive (predictand time series, see section \ref{sec:case_study_data}\ref{sec:data}). Based on these analog observations, the conditional empirical distribution considered as the probabilistic prediction for the target day (the day one wants to predict) is built.

Predictors for precipitation prediction, can be varied: e.g., geopotential height at different pressure levels and temporal windows, humidity variables, etc \citep[see][for a more detailed list of predictors]{BenDaoud2016}. The method often contains several levels of analogy, which are successive subsampling on predictors of different nature (e.g., atmospheric circulation, moisture variables, vertical motion, and air temperature).

The basis of the AM implementation considered here is the following \citep[same approach as][]{Guilbaud1997, Bontron2005, Marty2012, Horton2012, Radanovics2013, Chardon2014, Dayon2015, BenDaoud2016}:

\begin{enumerate}
	\item Preselection: to cope with seasonal effects, $n_{0}$ candidate dates are extracted from the archive within a period of four months centered around the target date, for every year of the archive. Alternatively, the candidate dates can be selected based on similar air temperature \citep[see][]{BenDaoud2016}.

	\item First level of analogy: $n_{1}$ dates are selected out of the preselected $n_{0}$, by means of an analogy ranking. The first level of analogy for precipitation prediction is often based on the atmospheric circulation. The similarity of the atmospheric circulation of the target date with the candidate situations is assessed on the geopotential height (at specific pressure levels, such as 500 and 1000~hPa, and over a certain spatial window, e.g., 12~h and 24~h) by means of the S1 criteria \citep[Eq. (\ref{eq:S1}), ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a defined spatial window. The S1 criteria is found more relevant than Euclidean distances for pressure fields by different studies \citep{Wilson1980, Woodcock1980, Guilbaud1998, Bontron2004}.

	\begin{equation}
	\label{eq:S1}
	S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
	{\displaystyle \sum_{i} max( \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert ) }
	\end{equation}
	where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points from the grid of the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values, the more similar the pressure fields.

	The $n_{1}$ dates with the lowest values of S1 are considered as analog situations, in terms of the atmospheric circulation, to the target day. The analog number $n_{1}$ is a parameter to calibrate.

	\item Subsequent level(s) of analogy: beyond the similarity of airflows, one may look for analogies in other variables of interest, such as moisture variables. Therefore, the $n_{1}$ analogs are subsampled once again on the basis of another variable, in order to obtain a lower number of analog dates, $n_{2}$. When the predictor is not a pressure field, the analogy criteria usually represents absolute distances: Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE), the latest being most often used.

	A second level of analogy based on thermodynamic variables was introduced by \citet{Vallee1986}, and \citet{Gibergans-Baguena2007}. After a systematic assessment of variables, \citet{Bontron2004} pointed out that a moisture index made of the product of relative humidity at 850~hPa and total precipitable water gives the best performance. This index does not represent an actual physical quantity, but expresses the water content of the air column and its proximity to saturation.

	This process can be repeated, by subsampling a decreasing number of analog dates, $n_{i}$, according to various meteorological variables.

	\item Probabilistic prediction: then, the daily observed precipitation amounts of the $n_{i}$ resulting dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day.

@@

The parameters to be defined (manually or automatically), for every level of analogy, are the following:

\begin{itemize}
	\item The selection of meteorological variables (predictors) that contain synoptic scale information having direct or indirect dependency with the target predictand.
	\item The pressure level, or height, at which the predictor is selected.
	\item The temporal window is the hour(s) of the day at which the predictors are considered.
	\item The spatial window is the domain on which predictors are compared. The ideal size of this area is the one that maximizes useful information and minimizes noise.
	\item The analogy criterion, needed to compare the variables on the chosen spatial and temporal windows, is a distance measure used for ranking observed situations according to their degree of similarity with the target situation.
	\item Eventual weights between the predictors \cite[e.g.,][]{Horton2012a, Junk2015}
	\item The ideal number of analog situations, which is the best compromise in order to take into account local variability and to maximize useful synoptic information \citep{Bontron2004}.
\end{itemize}

In order to calibrate the method, the Continuous Ranked Probability Score \citep[CRPS,][]{Brown1974, Matheson1976, Hersbach2000} is often used to assess AMs performances \citep[see, e.g.,][]{Bontron2004, Bontron2005, BenDaoud2008, Horton2012, Marty2012, Radanovics2013, Chardon2014, Junk2015, BenDaoud2016, Caillouet2016}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $l$ can be written:

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{l} \sum_{i=1}^{l} \left\{  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right\} 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value one otherwise. The mean CRPS is averaged on the calibration, respectively the validation periods.

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution of daily precipitation as the reference. The CRPSS (\textit{Continuous Ranked Probability Skill Score}) is thus defined as following:

\begin{equation}
\label{eq:CRPSS}
CRPSS = 1-\frac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the score value for the reference. An increase in CRPSS means a better prediction.

\section{Assessed Genetic Algorithms variants}
\label{sec:gas}

Genetic Algorithms \citep[GAs,][]{Holland1992b, Goldberg1989} are part of the family of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, inspired by some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. Unlike linear or local optimizations, GAs seek the global optimum on complex surfaces, theoretically without restriction, but with no guarantee to reach it \citep{Haupt2004}. The objective function to optimize (often named fitness function in this context) can be of different types \citep{Joines1996a}, but GAs must be adapted in order to perform optimally.

A key element of GAs configuration is finding the right balance between exploration and exploitation \citep{Back1992a, Smith1997a}. Exploration is characterized by a relatively high probability to assess the regions of the parameter space that have not yet been visited. This probability must be sufficiently large at the beginning of the optimization, so that the algorithm is capable of identifying the region where the global optimum is likely located. Exploitation is characterized by a local search in an area of interest, and generally makes small movements. The latter aims at finding small improvements at the end of the optimization.

@@

GAs optimize a population of $N$ individuals. Each individual contains a chromosome, which is here a vector of the AM parameters. Genes are the individual parameters constituting the chromosome. They can be either categorical (e.g., choice of the meteorological variable), discrete (e.g., number of analog dates to select), or continuous.

>>>>

\abstract{Analog methods are based on a statistical relationship between synoptic meteorological variables (predictors) and local weather (predictand, to be predicted). This relationship is defined by several parameters, which are often calibrated by means of a semi-automatic sequential procedure. This calibration approach is fast, but has strong limitations. It proceeds through successive steps, and thus cannot handle all parameters dependencies. Furthermore, it cannot automatically optimize some parameters, such as the selection of pressure levels and temporal windows (hours of the day) at which the predictors are compared.
In order to overcome these limitations, the global optimization technique of genetic algorithms is considered, which can jointly optimize all parameters of the method, and get closer to a global optimum, by taking into account the dependencies of the parameters. Moreover, it can objectively calibrate parameters that were previously assessed manually, and can take into account new degrees of freedom.
However, genetic algorithms must be tailored to the problem under consideration. Multiple combinations of algorithms were assessed, and new algorithms were developed (e.g., the \textit{chromosome of adaptive search radius}, which is found to be very robust), in order to provide recommendations regarding the use of genetic algorithms for optimizing several variants of analog methods. A global optimization approach provides new perspectives for the improvement of analog methods, and for their application to new regions or new predictands.
}

@@

Analog methods (AMs) rely on the hypothesis that similar situations, in terms of atmospheric circulation, are likely to result in similar local weather conditions \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. These methods consist of sampling a certain number of past situations, based on different synoptic-scale meteorological variables (predictors), in order to construct a probabilistic prediction for a local weather variable of interest (predictand). Some common usages of AMs are for operational precipitation forecasting \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015b, BenDaoud2016}, or more recently for precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. However, AMs or equivalent methods are also employed to predict temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

Although the method is rather simple, it contains several parameters to be optimized, such as the choice of the predictor variables, pressure levels and temporal windows (hours of the day) at which the predictors are compared along with the spatial domains used for the comparison, the analogy criteria (distance measure), and finally the number of analog situations to retain.

AMs must be adapted to every new considered region because the leading meteorological influences may be location-specific. Even the selection of the pressure levels and the temporal windows should be reconsidered, if not the predictor variable itself. Thus, before being applied in a forecasting or downscaling context, AMs must be calibrated for the given region, which is performed here in the perfect prognosis framework \citep{Klein1963}, in line with the majority of previously cited AM calibration procedures for precipitation predictions \cite[with the exception of][]{Hamill2006,Hamill2015b}.

A common approach to optimizing this method is by means of a semi-automatic sequential calibration procedure, which was developed by \citet{Bontron2004}, and is also described in \citet{BenDaoud2016} and extended by \citet{Radanovics2013}. This one determines some parameters of the method sequentially for each consecutive analogy level (e.g., on the atmospheric circulation or on a moisture index). It begins with a manual selection of the meteorological variables (e.g., geopotential height and relative humidity), the pressure levels, the temporal windows, and the initial analog numbers. Then, the spatial window over which the predictors are compared is optimized through an iterative growth of the domain, and the number of analogs is finally reassessed. A successive level of analogy can then be introduced, and its spatial window be optimized. The parameters of the preceding levels of analogy are not reassessed, except for the number of analog situations to preserve.

Thus, the sequential calibration procedure allows the optimization of a limited number of parameters (spatial windows and analog numbers), but the selection of predictor variables, pressure levels, and temporal windows must still be made manually. Testing multiple combinations of these presents a very combinatorial problem, which quickly becomes cumbersome, especially when considering multiple predictors within the same level of analogy. Thus, optimizing AMs by means of this sequential technique is laborious if little knowledge is available regarding the predictor-predictand relationship or the leading meteorological influences. Indeed, many combinations of parameters (predictor variables, pressure levels, and temporal windows) must be assessed. Moreover, proceeding to the optimization sequentially ignores potential dependencies between the parameters of the method, whether they are within a single level of analogy or between multiple levels, which could lead to another configuration if the parameters were calibrated together. Thus, there is a high risk of ending in a local optimum, and this cannot be avoided.

When creating this sequential calibration procedure, \citet{Bontron2004} was aware of the problem of dependencies between parameters, and wrote: ''\textit{We perceive here the combinatorial aspect of our problem: variables and spatial windows are not independent. We will present our results by first searching the best variable (e.g., selection of the pressure level and the temporal window for the geopotential height) on a chosen spatial window, and next, the best window for the chosen variable. However, even by repeating the process, are we sure to obtain the optimal combination?}'' Later in his work, he also wrote: ''\textit{Our approach, which is again to vary the parameters one by one -- the others being fixed in a more or less arbitrary manner -- may therefore not exactly lead us to the optimal solution}.'' \citet{Bliefernicht2010} has also confronted the combinatorial issue for the parameters of AMs, and concluded that one must be an expert in order to have a sense of their respective influence, sensitivity, and nonlinear interactions. When calibrating an AM, \citet{BenDaoud2010} also stated that ''\textit{the combinatory aspect related to the calibration was found to be too high for all the parameters to be calibrated simultaneously}.'' The simultaneous calibration of all parameters has not been undertaken so far.

Another optimization strategy, proposed by \citet{Junk2015}, allows for an automatic calibration of weights applied to the different predictors when processing the analogy criteria (distance function). Their strategy consists of a brute-force assessment of all possible combinations. This approach is possible in their implementation, because predictors are considered at a unique point (interpolated to the location of interest), at fixed hours, and at preselected pressure levels, leaving only the weights to be optimized. In the presently employed AM (described in section \ref{sec:am}), the number of parameters to optimize makes it impossible to proceed with a brute-force strategy.

In order to overcome these limitations, two optimization techniques have been assessed. First, \citet{Horton2012a} assessed the ability of the method of \citet{Nelder1965a}, based on a simplex algorithm. This technique did not provide satisfying results, and failed to converge toward a unique solution. The parameter space of the AMs can be very complex, and is inappropriate for a linear optimization technique. Thus, global optimization techniques are likely to be necessary in order to calibrate most AM variants, as they can optimize all parameters of all analogy levels simultaneously. In addition, they can avoid the systematic manual assessments of all pressure levels and temporal windows. Finally, they allow the testing of new degrees of freedom in AMs. The relevance of genetic algorithms (GAs) is demonstrated here, which does not exclude that other global optimization techniques could also work. Although using GAs to optimize AMs may be computationally intensive, once an AM is calibrated its employment in real-time operations or climate downscaling is very fast and lightweight.

This paper specifically describes how GAs should be used in order to successfully optimize several AMs. Indeed, the variants of GAs are numerous, and always need to be tailored to the addressed problem. This requires intensive and systematic comparisons of operators and options, in order to identify the key factors influencing the optimization, and the respective sensitivity of the options. Such analyses are presented here, resulting in recommendations for the use of GAs to optimize several AM implementations. The in-depth analysis of the benefits of such an approach in a specific case study will be the topic of a forthcoming paper.

This document begins by presenting the area of the case study and the relevant data (section \ref{sec:case_study_data}), the considered AM variants (section \ref{sec:am}), and the assessed GA options (section \ref{sec:gas}). Comparative analyses of these options are presented in section \ref{sec:assessment}, leading to the recommendations formulated in section \ref{sec:use}, and finally the conclusions presented in section \ref{sec:conclusions}.

@@

The area of the study is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l., and the area is 5524~km$^{2}$. Based on various climatological analyses \cite[see][for the details]{Horton2012a}, the gauging stations in the catchment were clustered in 10~subregions (Fig.\ \ref{fig:map}).

This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project, which aims to provide real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. AMs are used to provide real-time probabilistic precipitation forecasts, based on NWP (Numerical Weather Prediction) models outputs.

@@

AMs rely on two types of data: predictors, which are meteorological variables describing the state of the atmosphere at a synoptic scale, and the predictand, which is the local weather time series that is to be predicted.

When working in the perfect prognosis framework \citep{Klein1963}, the meteorological archive from which the predictors are extracted is usually a reanalysis dataset. Conversely, other applications of AMs for wind forecasting are based on a Model Output Statistics framework \citep[MOS, see][]{Glahn1972}, and thus employ forecast archives or reforecast products \citep[e.g.,][]{DelleMonache2013, DelleMonache2011, Alessandrini2015, Junk2015, Junk2015c}.

In the present study, the NCEP-NCAR reanalysis I dataset \citep[six-hourly, 17 pressure levels at a resolution of 2.5\degree, see][]{Kalnay1996} is employed, but this could be replaced with any other reanalysis dataset. This dataset is relatively old, although it is still widely used, and better results may be expected with more recent datasets. However, we can safely assume that if an optimization technique works for this reanalysis dataset, it will also work for an improved one.

Here, the predictand (which is to be predicted) is the daily precipitation (0540~UTC to 0540~UTC the following day) measured at the MeteoSwiss network of stations in the catchment of interest. The time series from every available gauging station were averaged over subregions of approximately 500~km$^{2}$ \citep[see][for details]{Horton2012a} in order to smooth local effects \citep{Obled2002, Marty2012}. The time series must be split into calibration and independent validation periods.

@@

AMs are based on the principle that two similar synoptic situations may result in similar local effects \citep{Lorenz1956, Lorenz1969}. Thus, they consist in searching for a certain number of past situations in a meteorological archive that present similar properties for chosen predictors, according to an analogy criteria, in order to extract the observed values of the local weather variable of interest from another archive (predictand time series, see section \ref{sec:case_study_data}\ref{sec:data}). Based on these analog observations, the conditional empirical distribution to be considered as the probabilistic prediction for the target day (the day one wishes to predict) is constructed.

Predictors for precipitation predictions can be varied. For example, the geopotential height at different pressure levels and temporal windows, or humidity variables \citep[see][for a more detailed list of predictors]{BenDaoud2016}. This method often contains several levels of analogy, which constitute of successive subsamplings on predictors of different natures (e.g., atmospheric circulation, moisture variables, vertical motion, and air temperature).

The basis of the AM implementation considered here is the following \citep[the same approach as][]{Guilbaud1997, Bontron2005, Marty2012, Horton2012, Radanovics2013, Chardon2014, Dayon2015, BenDaoud2016}:

\begin{enumerate}
	\item Preselection: To cope with seasonal effects, $n_{0}$ candidate dates are extracted from the archive within a period of four months centered around the target date, for every year of the archive. Alternatively, the candidate dates can be selected based on similar air temperatures \citep[see][]{BenDaoud2016}.

	\item First level of analogy: $n_{1}$ dates are selected out of the preselected $n_{0}$, by means of an analogy ranking. The first level of analogy for precipitation prediction is often based on the atmospheric circulation. The similarity between the atmospheric circulation of the target date and the candidate situations is assessed based on the geopotential height (at specific pressure levels, such as 500 and 1000~hPa, and over a certain spatial window, e.g., 12~h and 24~h) by means of the S1 criteria \citep[Eq. (\ref{eq:S1}), ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients over a defined spatial window. Various studies have found the S1 criteria to be more relevant than Euclidean distances for pressure fields \citep{Wilson1980, Woodcock1980, Guilbaud1998, Bontron2004}.

	\begin{equation}
	\label{eq:S1}
	S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
	{\displaystyle \sum_{i} max( \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert ) }
	\end{equation}
	where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points from the grid of the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values, the more similar the pressure fields are.

	The $n_{1}$ dates with the lowest S1 values are considered as analog situations, in terms of the atmospheric circulation, for the target day. The analog number $n_{1}$ is a parameter to be calibrated.

	\item Subsequent level(s) of analogy: Beyond the similarity of airflows, one may look for analogies in other variables of interest, such as moisture variables. Therefore, the $n_{1}$ analogs are subsampled once more on the basis of another variable, in order to obtain a lower number of analog dates, $n_{2}$. When the predictor is not a pressure field, the analogy criteria usually represents absolute distances, such as mean absolute error (MAE) or root mean squared error (RMSE), with the latter being most often employed.

	A second level of analogy, based on thermodynamic variables, was introduced by \citet{Vallee1986} and \citet{Gibergans-Baguena2007}. After a systematic assessment of variables, \citet{Bontron2004} pointed out that a moisture index consisting of the product of relative humidity at 850~hPa and total precipitable water achieves the best performance. This index does not represent an actual physical quantity, but expresses the water content of the air column and its proximity to saturation.

	This process can be repeated by subsampling a decreasing number of analog dates, $n_{i}$, according to various meteorological variables.

	\item Probabilistic prediction: Then, the daily observed amounts of precipitation for the $n_{i}$ resulting dates provide the empirical conditional distribution, considered as the probabilistic forecast for the target day.

@@

The parameters to be defined (manually or automatically), for every level of analogy, are:

\begin{itemize}
	\item The selection of meteorological variables (predictors), containing synoptic scale information, and having a direct or indirect dependency with the target predictand.
	\item The pressure level, or height, at which the predictor is selected.
	\item The temporal window is the hour(s) of the day at which the predictors are considered.
	\item The spatial window is the domain on which predictors are compared. The ideal size of this area is that which maximizes the useful information and minimizes noise.
	\item The analogy criterion, needed to compare the variables on the chosen spatial and temporal windows, is a distance measure, used to rank observed situations according to their degree of similarity with the target situation.
	\item Eventual weights between the predictors \cite[e.g.,][]{Horton2012a, Junk2015}.
	\item The ideal number of analog situations, which is the best compromise in order to take into account local variability and maximize useful synoptic information \citep{Bontron2004}.
\end{itemize}

In order to calibrate the method, the continuous ranked probability score \citep[CRPS,][]{Brown1974, Matheson1976, Hersbach2000} is often employed to assess the performances of AMs \citep[see, e.g.,][]{Bontron2004, Bontron2005, BenDaoud2008, Horton2012, Marty2012, Radanovics2013, Chardon2014, Junk2015, BenDaoud2016, Caillouet2016}. This allows the evaluation of predicted cumulative distribution functions $F(y)$, e.g., of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $l$ can be written as follows:

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{l} \sum_{i=1}^{l} \left\{  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right\} 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function, which is zero when $y-y_{i}^{0}<0$ and has a value of one otherwise. The mean CRPS is averaged over the calibration or the validation period.

In order to compare the value of the score with regard to a reference, one often considers its skill score expression, and employs the climatological distribution of daily precipitation as the reference. The CRPSS (\textit{continuous ranked probability skill score}) is thus defined as follows:

\begin{equation}
\label{eq:CRPSS}
CRPSS = 1-\frac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the score value for the reference. An increase in the CRPSS indicates a better prediction.

\section{Assessed genetic algorithm variants}
\label{sec:gas}

Genetic algorithms \citep[GAs,][]{Holland1992b, Goldberg1989} are part of the family of evolutionary algorithms \citep{Back1993b, Schwefel1993}, which are inspired by some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. Unlike linear or local optimizations, GAs seek the global optimum on complex surfaces, theoretically without restriction, but with no guarantee of reaching it \citep{Haupt2004}. The objective function to be optimized (often named the fitness function in this context) can be of various types \citep{Joines1996a}, but GAs must be adapted in order to perform optimally.

A key element of the configuration of GAs is to determine the correct balance between exploration and exploitation \citep{Back1992a, Smith1997a}. Exploration is characterized by a relatively high probability of assessing the regions of the parameter space that have not yet been visited. This probability must be sufficiently large at the beginning of the optimization, so that the algorithm is capable of identifying the region where the global optimum is likely to be located. Exploitation is characterized by a local search in an area of interest, and generally makes small movements. The latter aims at finding small improvements at the end of the optimization.

@@

GAs optimize a population of $N$ individuals. Each individual contains a chromosome, which here is a vector of the AM parameters. Genes are the individual parameters constituting the chromosome. These can be categorical (e.g., choice of the meteorological variable), discrete (e.g., number of analog dates to select), or continuous.
